03-modeling

### 2018-05-11 01_writing/Schratz_hyperspectral.tex
11/05/2018

\journal{Journal "Remote Sensing of environment"}

@@

	% geringerer Zeilenabstand
	%\setlength{\itemsep}{-\parsep}
	\acro{ANN}{Artificial Neural Network}
	\acro{AUROC}{Area Under the Receiver Operating Characteristics Curve}
	\acro{BRT}{Boosted Regression Trees}
	\acro{CART}{Classification and Regression Trees}
	\acro{CV}{cross-validation}
	\acro{ENM}{Environmental Niche Modeling}
	\acro{FPR}{False Positive Rate}
	\acro{GAM}{Generalized Additive Model}
	\acro{GBM}{Gradient Boosting Machine}
	\acro{GLM}{Generalized Linear Model}
	\acro{ICGC}{Institut Cartografic i Geologic de Catalunya}
	\acro{IQR}{Interquartile Range}
	\acro{WKNN}{Weighted $k$-nearest neighbor}
	\acro{MARS}{Multivariate Adaptive Regression Splines}
	\acro{MEM}{Maximum Entropy Model}
	\acro{NRI}{Normalized Ratio Index}
	\acro{LOWESS}{Locally Weighted Scatter Plot Smoothing}
	\acro{PISR}{Potential Incoming Solar Radiation}
	\acro{RBF}{Radial Basis Function}
	\acro{RF}{Random Forest}
	\acro{SDM}{Species Distribution Modeling}
	\acro{SVM}{Support Vector Machines}
	\acro{TPR}{True Positive Rate}
\end{acronym}

@@

Penalized regression was used to account for the large amount of highly correlated predictor variables.
The aim was to find the indices that best explain defoliation within the plots.

Due to the amount of highly correlated predictors in the dataset, the independence assumption of the predictors is violated.
To compensate for that, penalized regression penalizes the coefficients.
This leads to a substantial decrease in variance and better predictive performance.
However, it also results in biased coefficients that cannot be used for statistical inference anymore.
The resulting coefficients can be seen as a measure of variable importance.
The penalization terms that have been used in this work are explained in the following sections.

\subsubsection{L1 Penalization}

\subsubsection{L2 Penalization}

\subsubsection{Elasticnet}

@@

The first step was to decide which penalization term works best for the given data.
A nested 10-fold spatial \ac{CV} was conducted on every single plot and for the merged dataset, we ran a plot-based \ac{CV} for both performance evaluation and hyperparameter tuning.
% show plot of spatial partitioning for both?
This means that three out of the 4 plots formed the training set while the test set consisted of the remaining plot.

Next, we applied the winning method (Ridge) again on the plots and the merged dataset to find the most important coefficients for every plot.
Hyperparameter tuning of

@@

>>>>

\journal{Journal "Remote Sensing of Environment"}

@@

	% geringerer Zeilenabstand
	%\setlength{\itemsep}{-\parsep}
	\acro{ANN}{Artificial Neural Network}
	\acro{AUROC}{Area Under the Receiver Operating Characteristics Curve}
	\acro{BRT}{Boosted Regression Trees}
	\acro{CART}{Classification and Regression Trees}
	\acro{CV}{cross-validation}
	\acro{ENM}{Environmental Niche Modeling}
	\acro{FPR}{False Positive Rate}
	\acro{GAM}{Generalized Additive Model}
	\acro{GBM}{Gradient Boosting Machine}
	\acro{GLM}{Generalized Linear Model}
	\acro{ICGC}{Institut Cartografic i Geologic de Catalunya}
	\acro{IQR}{Interquartile Range}
	\acro{WKNN}{Weighted $k$-nearest neighbor}
	\acro{MARS}{Multivariate Adaptive Regression Splines}
	\acro{MEM}{Maximum Entropy Model}
	\acro{NRI}{Normalized Ratio Index}
	\acro{OLS}{Ordinary Least Squares}
	\acro{LOWESS}{Locally Weighted Scatter Plot Smoothing}
	\acro{PISR}{Potential Incoming Solar Radiation}
	\acro{RBF}{Radial Basis Function}
	\acro{RF}{Random Forest}
	\acro{RMSE}{Root Mean Square Error}
	\acro{RR}{Ridge Regression}
	\acro{RSS}{Residual Sum of Squares}
	\acro{SDM}{Species Distribution Modeling}
	\acro{SVM}{Support Vector Machines}
	\acro{TPR}{True Positive Rate}
\end{acronym}

@@

The aim of this work was to find the indices that best explain defoliation within the plots.
We used penalized regression to account for the large amount of highly correlated predictor variables.
In a standard \ac{OLS} regression one of the assumptions is that the predictor variables should be independent when minimizing the \ac{RSS} \citep{Bare1981, Hastie2001}:

\begin{equation}
	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2}
\end{equation}

where $\beta_{0}$ is the intercept, $\beta_{i}$ the coefficient, $x_{ij}$ the predictor variable and $y_{i}$ the response variable.

If this assumption is violated, regression coefficients can be highly biased.
They can even show the wrong sign and are very sensitive to adding new independent variables or data points to the model.
These points reduce the robustness and performance of \ac{OLS} regression when dealing with multicollinearity.
One approach to overcome these limitations is to penalize the coefficients.
This method leads to a substantial decrease in variance and better predictive performance compared to \ac{OLS} regression.
However, it also sacrifies the assumption of unbiased coefficients.
Hence, the resulting coefficients cannot be used for statistical inference but should be interpreted as a measure of variable importance.

\subsubsection{The ridge penalty}

In \ac{RR} (also called $\ell_{2}$ penalization) the assumption of unbiased coefficients is given up in favor of higher predictive accuracy and reduced variance \citep{Hastie2001}.
Coefficients are standardized and penalized for their size.
When minimizing the \ac{RSS}, \ac{RR} adds a penalization term $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ to the equation:

\begin{equation}
	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p}\beta_{j}^{2}
\end{equation}

where $\lambda >= 0$ is a tuning parameter responsible for the magnitude of penalization.
To make the second term (usually referred to as the \textit{shrinkage penalty}) of this equation small, the coefficients $\beta_{j}$ need to become small.
Unlike to Lasso however, predictors are not removed from the final model and will always be $\beta_{j} >= 0$.
Hence, $\lambda$ has the effect of shrinking the coefficients when minimizing the \ac{RSS}.
For $\lambda = 0$, no penalization is done and standard \ac{OLS} applies \citep{James2013}.
The \textit{shrinkage penalty} is only applied to the coefficients and not to the intercept.
Also, while \ac{OLS} generates only one set of coefficient estimates, \ac{RR} will create multiple sets for every value of $\lambda$.
In summary, the advantage of \ac{RR} is based on the bias-variance tradeoff: For $\lambda\to\infty$, the flexibility of the fit is reduced leading to a decrease in variance of the coefficients but also introduces a (substantial) bias.
For $\lambda = 0$, the variance is high but coefficients are unbiased \citep{James2013}.

\subsubsection{The lasso penalty}

While the Ridge penalty $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ shrinks the coefficients, it does not exclude any predictor from the model by setting the coefficient to zero.
This is what the Lasso penalty (also called $\ell_{1}$) does: It works similar to a "best subset selection" approach by only keeping the most important variables in the model and reducing the coefficients of unimportant ones down to zero.
This is done by a different \textit{shrinkage penalty} $\lambda \sum_{j=1}^{p} \mid\beta_{j}\mid$ that is added to the \ac{OLS} term when minimizing the \ac{RSS}:

\begin{equation}
	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p} \mid\beta_{j}\mid
\end{equation}

Subsequently, Lasso is performing variable selection and results in a sparse model that has the advantage of a simplied interpretability compared to Ridge models.
However, coeffcients are still biased due to penalization \citep{Hastie2001, James2013}.

\subsubsection{The elasticnet penalty}

The idea of the \textit{elasticnet penalty} is to combine both penalties Lasso and Ridge and weight these by a parameter $\alpha$ that that needs to be optimized.
It was introduced by \cite{Zou2005} with the idea to combine the advantages of both penalty terms.
The \textit{elasticnet penalty} is written as

\begin{equation}
	\lambda \sum_{j=1}^{p} \alpha\beta_{j}^{2} + (1 - \alpha) \mid\beta_{j}\mid
\end{equation}

where $\alpha$ weights the contribution of either the Lasso or Ridge penalty.
For $\alpha = 0.5$ both are weighted equally.
When $\alpha = 0$ only Ridge is used and for $\alpha = 1$ only Lasso applies \citep{Hastie2001}.

@@

\subsubsection{Selecting the best penalty}

As the introduced penalties behave different depending on the characteristics of the dataset, we first conducted a nested 10-fold spatial \ac{CV} to find the best method among those three.
The comparison was applied to all of the four plots using \ac{RMSE} as the error measure.
For both the performance evaluation and hyperparameter tuning a spatial sampling using $k$-means clustering was used \citep{sperrorest}.

\subsubsection{Extracting the most important variables}

Next, we applied the winning method (\ac{RR}) on all plots and additionally to the merged dataset of all plots to find the most important coefficients of every setup.
Hyperparameter tuning of $\lambda$ was done using again a spatial sampling again for the single plots.
For the merged dataset we used a plot-based sampling approach: One plot serves as the test partition while the remaining plots form the training set.
This approach has the advantage of a spatially independent test partition but also comes with the limitation of estimating the performance on a single repetition only as the restriction to the plot-level leaves no opportunity to create multiple resampling instances.

\subsubsection{Linking the variables to plot characteristics}

To interpret the outcomes of the models on a plot level, we linked the winning variables of each to attributes of the dataset that describe the underlying plot characteristics.
For example, a plot which a low tree density might inherit more information from the bare ground in the calculdated indices while a plot with a very high tree density might in contrast contain information from multiple trees.

@@

\section{Results}

\subsection{Comparison of penalties}

% parameter limits
\begin{table}[b!]
\centering
\caption[t]{Predictive performances of lasso, ridge and elasticnet penalties on the single plots and the merged dataset using \ac{RMSE} as the error measure.}
\begingroup\footnotesize
\begin{tabular}{llll}
	\\
	Plot/Penalty & Lasso  & Ridge  & Elasticnet \\
	\hline
	Laukiz 1     & 146.33 & 87.76  & 135.58     \\
	Laukiz 2     & 146.33 & 87.76  & 135.58     \\
	Luiando      &        &        &            \\
	Oiartzun     &        & 106.65 &            \\
	All Plots    & 56.88  & 55.88  & 56.76      \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:penalty_comparison}
\end{table}

### 2018-05-11 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
11/05/2018

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.05, upper = 5.5))
```

@@

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.05, upper = 5525),
                          makeNumericParam("alpha", lower = 0, upper = 1))
```

>>>>

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.5))
```

@@

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 5525),
                          makeNumericParam("alpha", lower = 0, upper = 1))
```
